---
title: "multilevel-tightness-analyses-OTH"
format: html
---

```{r}
library(tidyverse)
```

```{r}
oth <- read_csv("../../minerva/ResourceStress/Upstream/DT-resource-stress-combined.csv") %>%
  select(ID, OTH_30) %>%
  mutate(
    OTH_30 = if_else(OTH_30 == 99, NA, OTH_30)
  )

hzcats <- read_csv("../../minerva/Hazards/Datasets/DT-hzcats-society-level.csv")

ds_hz <- full_join(oth, hzcats, by = "ID")

sh <- read_csv("../../2026-Sharing-Analyses/sharing-dataset-for-analysis.csv") %>%
  select(ID, OWC, ea033, fd_freq_30, nonfd_2_freq_30, nonfd_2.5_freq_30)

tl_long <- read_csv("../../2026-Tightness-Analyses/2026-updates/data/tl_long.csv")

ds_sharing_tl <- inner_join(sh, tl_long, by = c("OWC"))

ds_FA <- full_join(ds_hz, ds_sharing_tl, by = c("ID", "OWC"))
```

```{r}
# inferred absence based on OTH_30 value
ds_FA_OTH <- ds_FA %>%
  mutate(
    comb_OTH_disaster_h9a_sev_exposure_30 = if_else(is.na(disaster_h9a_sev_exposure_30) & OTH_30 == 1, 0, disaster_h9a_sev_exposure_30),
    comb_OTH_allhazards_H9a_sev_exposure_30 = if_else(is.na(allhazards_H9a_sev_exposure_30) & OTH_30 == 1, 0, allhazards_H9a_sev_exposure_30),
    comb_OTH_fd3_est_count_30 = if_else(is.na(fd3_est_count_30) & OTH_30 == 1, 0, fd3_est_count_30),
    comb_OTH_fd_freq_30 = if_else(is.na(fd_freq_30) & OTH_30 == 1, 1, fd_freq_30),
    normative_ind_IA = factor(if_else(normative_est_count_30_IA > 0, 1, 0)),
  )
```

# TL5 Analyses: Overall gestalt level of TL in a society

## Impacts of food destroying hazard impacts on overall tightness (TL5)

### food destroying hazard frequency (a hazard must have H.9.a. \>= 3 to be counted for each society)

```{r}
tl5_mixed_fd_freq <- lmer(
  TL5 ~ comb_OTH_fd_freq_30 + ea033 +
    (1 | ID),
  data = ds_FA_OTH
)

summary(tl5_mixed_fd_freq)
```

ID groups vary around the central mean average of TL5 by about 0.243 units, and the standard deviation of this is about 0.4926.

More frequent food-destroying hazards are associated with lower TL5 scores (β = −0.082, p = .033), while greater political hierarchy is associated with higher TL5 scores (β = 0.150, p = .003), after accounting for repeated observations within societies.

### subdomains

```{r}
domain_nums <- sort(unique(ds_FA_OTH$Domain_name))

for (d in domain_nums) {
  cat("\n\n============================\n")
  cat("Domain:", d, "\n")
  cat("============================\n\n")
  
  domain_data <- ds_FA_OTH %>%
    filter(Domain_name == d)
  
  model <- lm(
  TL5 ~ comb_OTH_fd_freq_30 + ea033,
  data = domain_data
)
  
  print(summary(model))
}
```

| TL5 domain | Hazard β | p-value | Interpretation |
|------------------|-----------------:|-----------------:|------------------|
| Funeral | −0.091 | .054 | Negative, borderline/marginal evidence |
| Gender | **−0.126** | **.025** | **Significant negative association** |
| Law | −0.027 | .587 | Very little evidence of an association |
| Marriage | −0.086 | .080 | Negative, suggestive but not significant |
| Sex | −0.078 | .149 | Negative, but uncertain |
| Social | −0.113 | .105 | Negative, but uncertain |
| **Overall TL5** | **−0.082** | **.033** | **Significant overall negative association** |

### food destroying weight

```{r}
tl5_mixed_weighted <- lmer(
  TL5 ~ comb_OTH_disaster_h9a_sev_exposure_30 + ea033 +
    (1 | ID),
  data = ds_FA_OTH
)

summary(tl5_mixed_weighted)
```

### Control for normative hazards

```{r}
tl5_mixed_fd_freq_norm <- lmer(
  TL5 ~ comb_OTH_fd_freq_30 + ea033 + normative_ind_IA +
    (1 | ID),
  data = ds_FA_OTH
)

summary(tl5_mixed_fd_freq_norm)
```

```{r}
domain_nums <- sort(unique(ds_FA_OTH$Domain_name))

for (d in domain_nums) {
  cat("\n\n============================\n")
  cat("Domain:", d, "\n")
  cat("============================\n\n")
  
  domain_data <- ds_FA_OTH %>%
    filter(Domain_name == d)
  
  model <- lm(
  TL5 ~ comb_OTH_fd_freq_30 + ea033 + normative_ind_IA,
  data = domain_data
)
  
  print(summary(model))
}
```

Nothing...

# TL1 Analyses: Emphasis of strong rules for appropriate behavior across domains

## Food-destroying frequency

```{r}
tl1_mixed <- lmer(
  TL1 ~ comb_OTH_fd_freq_30 + ea033 +
    (1 | ID),
  data = ds_FA_OTH
)

summary(tl1_mixed)
```

Food-destroying hazard frequency was negatively associated with emphasis on strong rules for appropriate behavior, $\beta_{fdfreq} = −0.102$, SE = 0.038, p = .010, controlling for political hierarchy. The model included 458 domain-level observations nested within 85 societies.

```{r}
domain_nums <- sort(unique(ds_FA_OTH$Domain_name))

for (d in domain_nums) {
  cat("\n\n============================\n")
  cat("Domain:", d, "\n")
  cat("============================\n\n")
  
  domain_data <- ds_FA_OTH %>%
    filter(Domain_name == d)
  
  model <- lm(
  TL1 ~ comb_OTH_fd_freq_30 + ea033,
  data = domain_data
)
  
  print(summary(model))
}
```

### Control for normative hazards

```{r}
tl1_mixed_norm <- lmer(
  TL1 ~ comb_OTH_fd_freq_30 + ea033 + normative_ind_IA + 
    (1 | ID),
  data = ds_FA_OTH
)

summary(tl1_mixed_norm)
```

```{r}
domain_nums <- sort(unique(ds_FA_OTH$Domain_name))

for (d in domain_nums) {
  cat("\n\n============================\n")
  cat("Domain:", d, "\n")
  cat("============================\n\n")
  
  domain_data <- ds_FA_OTH %>%
    filter(Domain_name == d)
  
  model <- lm(
  TL1 ~ comb_OTH_fd_freq_30 + ea033 + normative_ind_IA,
  data = domain_data
)
  
  print(summary(model))
}
```



# Controlling for Irrigation


Following discussion at the 9/16 Minerva meeting