---
title: "oth-hazard-exploration"
format: html
---

```{r}
library(tidyverse)

oth <- read_csv("../../minerva/ResourceStress/Upstream/DT-resource-stress-combined.csv") %>%
  select(ID, OTH_30) %>%
  mutate(
    OTH_30 = if_else(OTH_30 == 99, NA, OTH_30)
  )

hzcats <- read_csv("../../minerva/Hazards/Datasets/DT-hzcats-society-level.csv")

# there are roughly 10 cases coded NA for OTH_30
inner_join(oth, hzcats, by = "ID") %>%
  filter(!is.na(OTH_30))
```

```{r}
# create combined dataset
ds_hz <- full_join(oth, hzcats, by = "ID")

sh <- read_csv("../../2026-Sharing-Analyses/sharing-dataset-for-analysis.csv") %>%
  select(ID, OWC, ea033, fd_freq_30, nonfd_2_freq_30, nonfd_2.5_freq_30)

tl_long <- read_csv("../../2026-Tightness-Analyses/2026-updates/data/tl_long.csv")

ds_sharing_tl <- inner_join(sh, tl_long, by = c("OWC"))

ds_FA <- full_join(ds_hz, ds_sharing_tl, by = c("ID", "OWC"))
```

```{r}
# inferred absence based on OTH_30 value
ds_FA_OTH <- ds_FA %>%
  mutate(
    disaster_h9a_sev_exposure_30_OTH = if_else(is.na(disaster_h9a_sev_exposure_30) & OTH_30 == 1, 0, disaster_h9a_sev_exposure_30),
    fd3_est_count_30_OTH = if_else(is.na(fd3_est_count_30) & OTH_30 == 1, 0, fd3_est_count_30),
    fd_freq_30_OTH = if_else(is.na(fd_freq_30) & OTH_30 == 1, 1, fd_freq_30),
    # nonfd_2_freq_30 = if_else(is.na(nonfd_2_freq_30) & OTH_30 == 1, 1, NA),
    # nonfd_2.5_freq_30 = if_else(is.na(nonfd_2.5_freq_30) & OTH_30 == 1, 1, NA)
     normative_ind_IA = factor(if_else(normative_est_count_30_IA > 0, 1, 0)),
  )
```

# TL5 Analyses: Overall gestalt level of TL in a society

```{r}
library(lme4)
library(lmerTest) # gives p-values
```

## Food-destroying Freq

```{r}
tl5_mixed_fd_freq <- lmer(
  TL5 ~ fd_freq_30_OTH + ea033 +
    (1 | ID),
  data = ds_FA_OTH
)

summary(tl5_mixed_fd_freq)
```

Food-destroying hazard frequency was negatively associated with TL5 scores ($\beta_{fdfreq} = −0.082$ , SE = 0.038, p = 0.033), such that greater food destroying hazard frequency was associated with lower TL5 scores after controlling for political hierarchy. Political hierarchy was positively associated with TL5 (β = 0.150, SE = 0.049, p = .003, N = 452 observations; n = 85 societies).

```{r}
domain_nums <- sort(unique(ds_FA_OTH$Domain_name))

for (d in domain_nums) {
  cat("\n\n============================\n")
  cat("Domain:", d, "\n")
  cat("============================\n\n")
  
  domain_data <- ds_FA_OTH %>%
    filter(Domain_name == d)
  
  model <- lm(
  TL5 ~ fd_freq_30_OTH + ea033,
  data = domain_data
)
  
  print(summary(model))
}
```

```{r}
data_gender <- ds_FA_OTH %>%
    filter(Domain_name == "gender")

plot(data_gender$fd_freq_30_OTH, data_gender$TL5, )

ggplot(
  data_gender,
  aes(
    x = TL5,
    y =fd_freq_30_OTH 
  )
) +
  geom_smooth(method = "loess",se = TRUE) +
  geom_jitter(
    width = 0.1,
    height = 0,
    alpha = 0.7
  )
```

## Food destroying disasters

```{r}
tl5_mixed_fd_dis <- lmer(
  TL5 ~ fd3_est_count_30_OTH + ea033 +
    (1 | ID),
  data = ds_FA_OTH
)

summary(tl5_mixed_fd_dis)
```

# TL1 Analyses: Emphasis of strong rules for appropriate behavior across domains

## Food-destroying frequency

```{r}
tl1_mixed <- lmer(
  TL1 ~ fd_freq_30_OTH + ea033 +
    (1 | ID),
  data = ds_FA_OTH
)

summary(tl1_mixed)
```

Food-destroying hazard frequency was negatively associated with emphasis on strong rules for appropriate behavior, $\beta_{fdfreq} = −0.102$, SE = 0.038, p = .010, controlling for political hierarchy. The model included 458 domain-level observations nested within 85 societies.

```{r}
domain_nums <- sort(unique(ds_FA_OTH$Domain_name))

for (d in domain_nums) {
  cat("\n\n============================\n")
  cat("Domain:", d, "\n")
  cat("============================\n\n")
  
  domain_data <- ds_FA_OTH %>%
    filter(Domain_name == d)
  
  model <- lm(
  TL1 ~ fd_freq_30_OTH + ea033,
  data = domain_data
)
  
  print(summary(model))
}
```

# TL3 Analyses: Punishment in violation of norms across domains

## Food-destroying frequency

```{r}
tl3_mixed <- lmer(
  TL3 ~ fd_freq_30_OTH + ea033 +
    (1 | ID),
  data = ds_FA_OTH
)

summary(tl3_mixed)
```

```{r}
domain_nums <- sort(unique(ds_FA_OTH$Domain_name))

for (d in domain_nums) {
  cat("\n\n============================\n")
  cat("Domain:", d, "\n")
  cat("============================\n\n")
  
  domain_data <- ds_FA_OTH %>%
    filter(Domain_name == d)
  
  model <- lm(
  TL3 ~ fd_freq_30_OTH + ea033,
  data = domain_data
)
  
  print(summary(model))
}
```
