---
title: "tightness-variables"
format: html
---

```{r}
library(tidyverse)

ds_FA <- read_csv("../../2026-Sharing-Analyses/sharing-dataset-for-analysis.csv")

tl <- read_csv("../../minerva/Tight-Loose/Datasets/DT-TL-FinalData.csv") %>%
  mutate(
    OWC = str_trim(OWC, side = c("both"))
  )

ds <- full_join(ds_FA, tl, by = c("OWC"))

hz_cat <- read_csv("../../minerva/Hazards/Datasets/DT-hzcats-society-level.csv") %>%
  select(ID, allhazards_est_count_30, disaster_est_count_30, disaster_est_count_30_IA)

ds <- inner_join(ds, hz_cat, by = c("ID")) %>%
  filter(ID != 73 & ID != 218)

```

# Evaluating the subdomains of cultural tightness as a means to be integrated within the 2026 PNAS project
It was made known that in evaluating overall cultural tightness, norms and rules were both expressed within this one variable. However, there is a reasonable concern that just because there may be rules the norm may not be consistent in following those rules or vice versa. This idea is what is driving the exploration of cultural tightness subdomains. 

There are 6 subdomains: Law and Ethics (General), Socialization, Gender, Marraige, Sexuality, and Funerals/Mourning

```{r}
tl_vars <- c(
  "General_TL1_avg_final", "General_TL2_avg_final", "General_TL3_avg_final",
  "General_TL4_avg_final", "General_TL5_avg_final", "General_TL5B_avg_final",
  "Soc_TL1_avg_final", "Soc_TL2_avg_final", "Soc_TL3_avg_final",
  "Soc_TL4_avg_final", "Soc_TL5_avg_final",
  "Gender_TL1_avg_final", "Gender_TL2_avg_final", "Gender_TL3_avg_final",
  "Gender_TL4_avg_final", "Gender_TL5_avg_final",
  "Mar_TL1_avg_final", "Mar_TL2_avg_final", "Mar_TL3_avg_final",
  "Mar_TL4_avg_final", "Mar_TL5_avg_final",
  "Sex_TL1_avg_final", "Sex_TL2_avg_final", "Sex_TL3_avg_final",
  "Sex_TL4_avg_final", "Sex_TL5_avg_final",
  "FM_TL1_avg_final", "FM_TL2_avg_final", "FM_TL3_avg_final",
  "FM_TL4_avg_final", "FM_TL5_avg_final",
  "General_TL1_pref_final", "General_TL2_pref_final", "General_TL3_pref_final",
  "General_TL4_pref_final", "General_TL5_pref_final", "General_TL5B_pref_final",
  "Soc_TL1_pref_final", "Soc_TL2_pref_final", "Soc_TL3_pref_final",
  "Soc_TL4_pref_final", "Soc_TL5_pref_final",
  "Gender_TL1_pref_final", "Gender_TL2_pref_final", "Gender_TL3_pref_final",
  "Gender_TL4_pref_final", "Gender_TL5_pref_final",
  "Mar_TL1_pref_final", "Mar_TL2_pref_final", "Mar_TL3_pref_final",
  "Mar_TL4_pref_final", "Mar_TL5_pref_final",
  "Sex_TL1_pref_final", "Sex_TL2_pref_final", "Sex_TL3_pref_final",
  "Sex_TL4_pref_final", "Sex_TL5_pref_final",
  "FM_TL1_pref_final", "FM_TL2_pref_final", "FM_TL3_pref_final",
  "FM_TL4_pref_final", "FM_TL5_pref_final",
  "TL_Ov_Avg", "TL_Ov_Avg_Pref"
)

ds_tl_numeric <- ds %>%
  mutate(across(all_of(tl_vars), ~ as.numeric(.)))

tl_coverage <- ds_tl_numeric %>%
  summarise(
    across(
      all_of(tl_vars),
      list(
        n_coverage = ~ sum(!is.na(.)),
        n_na = ~ sum(is.na(.)),
        min = ~ min(., na.rm = TRUE),
        max = ~ max(., na.rm = TRUE)
      ),
      .names = "{.col}_{.fn}"
    )
  ) %>%
  pivot_longer(
    everything(),
    names_to = c("variable", ".value"),
    names_pattern = "(.+)_(n_coverage|n_na|min|max)$"
  ) %>%
  mutate(
    range = paste0(min, "–", max)
  ) %>%
  select(variable, n_coverage, n_na, min, max, range)

tl_coverage
```

# TL1

## food destroying severity
```{r}
tl1_fd <- ds[complete.cases(ds[, c("disaster_h9a_sev_exposure_30", "General_TL1_avg_final")]), ]
ggplot(tl1_fd, aes(x = disaster_h9a_sev_exposure_30, y = as.numeric(General_TL1_avg_final) )) +
  geom_jitter(height = 0.05, alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE, color = "#66c2a4")
```

```{r}
ggplot(tl1, aes(
  x = factor(General_TL1_avg_final),
  y = allhazards_H9a_sev_exposure_30
)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.5) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.6) +
  labs(
    x = "General_TL1_avg_final",
    y = "allhazards_H9a_sev_exposure_30"
  )
```

## societies w/ no normative hazards
```{r}
tl1_2 <- tl1 %>%
  filter(normative_est_count_30_IA == 0)

ggplot(tl1_2, aes(x = as.numeric(General_TL1_avg_final), y = disaster_h9a_sev_exposure_30)) +
  geom_jitter(height = 0.05, alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE, color = "#66c2a4")

test <- lm(General_TL1_avg_final ~ allhazards_H9a_sev_exposure_30_IA, data = tl1_2)


# societies w/ 1 normative hazard
tl1_3 <- tl1 %>%
  filter(normative_est_count_30_IA > 0 & normative_est_count_30_IA <= 20)

ggplot(tl1_3, aes(x = as.numeric(General_TL1_avg_final), y = allhazards_H9a_sev_exposure_30)) +
  geom_jitter(height = 0.05, alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE, color = "#66c2a4")

test <- lm(General_TL1_avg_final ~ allhazards_H9a_sev_exposure_30, data = tl1_2)

test2 <- tl1_2 %>%
  select(ID, disaster_h9a_sev_exposure_30, General_TL1_avg_final, DQ)
```


```{r}
tl1 <- ds[complete.cases(ds[, c("allhazards_H9a_sev_exposure_30_IA", "General_TL1_avg_final")]), ]

ggplot(tl1, aes(x = allhazards_H9a_sev_exposure_30, y = as.numeric(General_TL1_avg_final) )) +
  geom_jitter(height = 0.05, alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE, color = "#66c2a4")
```

```{r}
ggplot(tl1, aes(
  x = factor(General_TL1_avg_final),
  y = allhazards_H9a_sev_exposure_30
)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.5) +
  geom_jitter(width = 0.1, height = 0, alpha = 0.6) +
  labs(
    x = "General_TL1_avg_final",
    y = "allhazards_H9a_sev_exposure_30"
  )
```


# TL3
## societies w/ no normative hazards
```{r}
tl3 <- ds[complete.cases(ds[, c("allhazards_H9a_sev_exposure_30", "General_TL3_avg_final")]), ]
ggplot(tl3, aes(x = allhazards_H9a_sev_exposure_30, y = as.numeric(General_TL3_avg_final) )) +
  geom_jitter(height = 0.05, alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE, color = "#66c2a4")

tl3_fd <- ds[complete.cases(ds[, c("disaster_h9a_sev_exposure_30_IA", "General_TL3_avg_final")]), ]
ggplot(tl3_fd, aes(x = disaster_h9a_sev_exposure_30_IA, y = as.numeric(General_TL3_avg_final) )) +
  geom_jitter(height = 0.05, alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE, color = "#66c2a4")

```

