---
title: "CA-long-format-analyses"
format: html
---

```{r}
library(tidyverse)

ca <- read_csv("collective-action-long-format.csv")

hz <- read_csv("../../minerva/Hazards/Datasets/DT-hzcats-society-level.csv")

sh <- read_csv("../../2026-Sharing-Analyses/sharing-dataset-for-analysis.csv")
sh <- sh %>% select(ID, allhazards_nonfd_sev_exposure2.5_30, allhazards_nonfd_sev_exposure2.5_30_IA, allhazards_nonfd_sev_exposure2_30, allhazards_nonfd_sev_exposure2_30_IA)

hz <- inner_join(hz, sh, by = c("ID"))
# remove outliers identified in cluster analysis  

hz <- hz %>% filter(ID != 73 & ID != 218)

hz_vars <- c(
  "normative_est_count_30",
  "normative_est_count_30_IA",
  "disaster_est_count_30",
  "disaster_est_count_30_IA",
  "allhazards_sev_exposure_30",
  "allhazards_sev_exposure_30_IA",
  "allhazards_H9a_sev_exposure_30",
  "allhazards_H9a_sev_exposure_30_IA",
  "allhazards_nonfd_sev_exposure2_30",
  "allhazards_nonfd_sev_exposure2_30_IA",
  "allhazards_nonfd_sev_exposure2.5_30",
  "allhazards_nonfd_sev_exposure2.5_30_IA"
)

hz_dat <- hz %>%
  select(ID, OWC, all_of(hz_vars)) %>%
  mutate(
    normative_ind = case_when(
      is.na(normative_est_count_30) ~ NA_integer_,
      normative_est_count_30 > 0 ~ 1L,
      TRUE ~ 0L
    ),
    normative_ind_IA = case_when(
      is.na(normative_est_count_30_IA) ~ NA_integer_,
      normative_est_count_30_IA > 0 ~ 1L,
      TRUE ~ 0L
    ),
    across(
      all_of(hz_vars),
      ~ as.numeric(scale(.x)),
      .names = "Z_{.col}"
    )
  )


```
```{r}
ca_hz <- ca %>%
  left_join(
    hz_dat,
    by = "OWC",
    relationship = "many-to-one"
  )
```

# Does the presence of collective action vary based on differing hazard environments?
mixed effects logistic model or ordinal mixed effects model (if wanting to use CAL1 directly)
cooperation_present ~ environmental_variation +
                      EA_dependence +
                      subsistence_domain +
                      (1 | society)
                      
Among subsistence types practiced by a society, is hazard exposure associated with whether cooperative subsistence activity occurs?
```{r}
library(lme4)

ca_presence <- ca_hz %>%
  filter(
    practiced == 1,
    !is.na(cooperation_present)
  ) %>%
  mutate(
     Z_EA_dependence = as.numeric(scale(EA_dependence))
  )

# all hazards severity
m_presence <- glmer(
  cooperation_present ~
    Z_allhazards_sev_exposure_30 +
    normative_ind_IA +
    Z_EA_dependence +
    factor(subsistence_domain) +
    (1 | OWC),
  data = ca_presence,
  family = binomial()
)

summary(m_presence)


```

                      

# when cooperation occurs, is there a connection between environmental impact and specific features of collective action

CAL1a ~ environmental_variation + EA_dependence + ...
CAL1b ~ environmental_variation + EA_dependence + ...
