---
title: "ca-variable-examination"
format: html
---

```{r}
library(tidyverse)

ca <- read_csv("../../minerva/CollectiveAction/DT-CollectiveAction_raw2025.csv") %>%
  mutate(
    OWC = str_trim(OWC, side = c("both"))
  )

ca <- ca[, -c(30:44)]

ca <- ca %>%
  mutate(across(everything(), ~ ifelse(.x %in% c(88, 99), NA, .x))) %>%
  rename(`GA-CAL1A-R` = `GA-CALA-R`)

```

```{r}
colnames(ca)
```
# Gathering
```{r}
# Gathering
ga <- ca %>%
  select(starts_with("GA"))

# across all gathering variables we want to see data coverage for each variable
ga_coverage <- ga %>%
  summarise(across(
    everything(),
    ~ sum(!is.na(.)),
    .names = "{.col}"
  )) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "non_missing_n"
  ) %>%
  mutate(
    total_n = nrow(ga),
    percent_coverage = 100 * non_missing_n / total_n
  )

ga_coverage
```

```{r}
table(ga$`GA-CAL1-R`)
table(ga$`GA-CALA-R`)
table(ga$`GA-CAL1B-R`)
table(ga$`GA-CAL1C-R`)
```

# Hunting
```{r}
hu <- ca %>%
  select(starts_with("HU"))

# across all gathering variables we want to see data coverage for each variable
hu_coverage <- hu %>%
  summarise(across(
    everything(),
    ~ sum(!is.na(.)),
    .names = "{.col}"
  )) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "non_missing_n"
  ) %>%
  mutate(
    total_n = nrow(hu),
    percent_coverage = 100 * non_missing_n / total_n
  )

hu_coverage
```

```{r}
table(hu$`HU-CAL1-R`)
table(hu$`HU-CAL1A-R`)
table(hu$`HU-CAL1B-R`)
table(hu$`HU-CAL1C-R`)
```
# Fishing
```{r}
fi <- ca %>%
  select(starts_with("FI"))

# across all gathering variables we want to see data coverage for each variable
fi_coverage <- fi %>%
  summarise(across(
    everything(),
    ~ sum(!is.na(.)),
    .names = "{.col}"
  )) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "non_missing_n"
  ) %>%
  mutate(
    total_n = nrow(fi),
    percent_coverage = 100 * non_missing_n / total_n
  )

fi_coverage
```

```{r}
table(fi$`FI-CAL1-R`)
table(fi$`FI-CAL1A-R`)
table(fi$`FI-CAL1B-R`)
table(fi$`FI-CAL1C-R`)
```

# Animal Husbandry
```{r}
ah <- ca %>%
  select(starts_with("AH"))

# across all gathering variables we want to see data coverage for each variable
ah_coverage <- ah %>%
  summarise(across(
    everything(),
    ~ sum(!is.na(.)),
    .names = "{.col}"
  )) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "non_missing_n"
  ) %>%
  mutate(
    total_n = nrow(ah),
    percent_coverage = 100 * non_missing_n / total_n
  )

ah_coverage
```

```{r}
table(ah$`AH-CAL1-R`)
table(ah$`AH-CAL1A-R`)
table(ah$`AH-CAL1B-R`)
table(ah$`AH-CAL1C-R`)
```

# Agriculture
```{r}
ag <- ca %>%
  select(starts_with("AG"))

# across all gathering variables we want to see data coverage for each variable
ag_coverage <- ag %>%
  summarise(across(
    everything(),
    ~ sum(!is.na(.)),
    .names = "{.col}"
  )) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "non_missing_n"
  ) %>%
  mutate(
    total_n = nrow(ag),
    percent_coverage = 100 * non_missing_n / total_n
  )

ag_coverage
```

```{r}
table(ag$`AG-CAL1-R`)
table(ag$`AG-CAL1A-R`)
table(ag$`AG-CAL1B-R`)
table(ag$`AG-CAL1C-R`)
```


# indicator domains based off of whether subsistence strategy is practiced (w/o reference to collective participation)

```{r}
new_ca <- ca %>%
  mutate(
    GA_practiced = if_else(`GA-CAL1-R` == 77, 0, 1),
    HU_practiced = if_else(`HU-CAL1-R` == 77, 0, 1),
    FI_practiced = if_else(`FI-CAL1-R` == 77, 0, 1),
    AH_practiced = if_else(`AH-CAL1-R` == 77, 0, 1),
    AG_practiced = if_else(`AG-CAL1-R` == 77, 0, 1),

    across(
      matches("CAL1A|CAL1B|CAL1"),
      ~ if_else(.x == 77, 0, as.numeric(.x))
    ),
    across(
      contains("CAL1B"),
      ~ case_when(
        .x == 0 ~ 0,
        .x == 4 ~ 1,
        .x == 3 ~ 2,
        .x == 2 ~ 3,
        .x == 1 ~ 4,
        TRUE ~ NA_real_
      ),
      .names = "{.col}_increasing"
    )
  )
```


## take a look individually at EA data coverage
```{r}
# GA 
table(ca$`EA-GA`)
table(ca$`EA-HU`)
table(ca$`EA-FI`)
table(ca$`EA-AH`)
table(ca$`EA-AG`)
```


# examine each society's coverage across EA and CA variable pairs
```{r}
pair_indicators <- new_ca %>%
  mutate(
    GA_pair = if_else(`EA-GA` > 4 & `GA-CAL1-R` >= 0, 1, 0),
    HU_pair = if_else(`EA-HU` > 4 & `HU-CAL1-R` >= 0, 1, 0),
    FI_pair = if_else(`EA-FI` > 4 & `FI-CAL1-R` >= 0, 1, 0),
    AH_pair = if_else(`EA-AH` > 4 & `AH-CAL1-R` >= 0, 1, 0),
    AG_pair = if_else(`EA-AG` > 4 & `AG-CAL1-R` >= 0, 1, 0),

    GA_ca_gt0 = if_else(`EA-GA` > 4 & `GA-CAL1-R` > 0, 1, 0),
    HU_ca_gt0 = if_else(`EA-HU` > 4 & `HU-CAL1-R` > 0, 1, 0),
    FI_ca_gt0 = if_else(`EA-FI` > 4 & `FI-CAL1-R` > 0, 1, 0),
    AH_ca_gt0 = if_else(`EA-AH` > 4 & `AH-CAL1-R` > 0, 1, 0),
    AG_ca_gt0 = if_else(`EA-AG` > 4 & `AG-CAL1-R` > 0, 1, 0)
  ) %>%
  mutate(
    pair_coverage_count = rowSums(across(ends_with("_pair")), na.rm = TRUE),
    ca_gt0_count = rowSums(across(ends_with("_ca_gt0")), na.rm = TRUE),
    any_domain_pair = if_else(pair_coverage_count > 0, 1, 0)
  )



table(pair_indicators$pair_coverage_count)

pair_summary <- pair_indicators %>%
  summarise(
    across(
      ends_with("_pair"),
      list(
        n = ~ sum(.x == 1, na.rm = TRUE),
        prop = ~ mean(.x == 1, na.rm = TRUE)
      )
    )
  ) %>%
  pivot_longer(
    everything(),
    names_to = c("domain", ".value"),
    names_pattern = "(.*)_pair_(n|prop)"
  )

pair_summary

ca_gt0_summary <- pair_indicators %>%
  summarise(
    across(
      ends_with("_ca_gt0"),
      ~ sum(.x == 1, na.rm = TRUE)
    )
  ) %>%
  pivot_longer(
    everything(),
    names_to = "domain",
    values_to = "n_pairs_w_collective_action"
  ) %>%
  mutate(
    domain = str_remove(domain, "_ca_gt0")
  )

ca_gt0_summary
```

```{r}
pair_indicators %>%
  ggplot(aes(x = pair_coverage_count)) +
  geom_bar() +
  labs(
    x = "Number of usable EA–CAL1 domain pairs",
    y = "Number of societies",
    title = "Coverage of subsistence-domain pairs (w/ EA > 4) across societies"
  ) +
  theme_minimal()

```
```{r}
ea_vars <- c("EA-GA", "EA-HU", "EA-FI", "EA-AH", "EA-AG")

cal1_vars <- c(
  "GA-CAL1-R",
  "HU-CAL1-R",
  "FI-CAL1-R",
  "AH-CAL1-R",
  "AG-CAL1-R"
)

ea_max_by_society <- new_ca %>%
   filter(if_any(all_of(cal1_vars), ~ !is.na(.x))) %>%
   mutate(
    max_EA_subsistence = pmax(
      `EA-GA`,
      `EA-HU`,
      `EA-FI`,
      `EA-AH`,
      `EA-AG`,
      na.rm = TRUE
    )
   )

ggplot(ea_max_by_society, aes(x = factor(max_EA_subsistence))) +
  geom_bar() +
  labs(
    x = "Max EA subsistence score across subsistence types",
    y = "Number of societies",
    title = "Spread of Maximum Subsistence Type Scores Across Societies"
  ) +
  theme_minimal()
```

# make new variable: ca_ea_pair_score

```{r}
hz <- read_csv("../../minerva/Hazards/Datasets/DT-hzcats-society-level.csv")

ds <- inner_join(hz, pair_indicators, by = c("ID" = "SCCS ID")) # something is wrong with the OWCs for the CA dataset since we're losing cases when we merge by it...will not be using to join

owcs <- anti_join(hz, pair_indicators, by = c("OWC")) # discrepancies between OWCs listed here
```

```{r}
ca_hz <- ds %>%
  filter(any_domain_pair == 1) %>%
  pivot_longer(
    cols = c(`EA-GA`, `EA-HU`, `EA-FI`, `EA-AH`, `EA-AG`),
    names_to = "domain",
    values_to = "EA_score"
  ) %>%
  mutate(
    domain = str_remove(domain, "EA-"),
    CAL1_score = case_when(
      domain == "GA" ~ `GA-CAL1-R`,
      domain == "HU" ~ `HU-CAL1-R`,
      domain == "FI" ~ `FI-CAL1-R`,
      domain == "AH" ~ `AH-CAL1-R`,
      domain == "AG" ~ `AG-CAL1-R`,
      TRUE ~ NA_real_
    )
  ) %>%
  filter(EA_score > 4, CAL1_score >= 0) %>%
  group_by(ID) %>%
  # slice_max(EA_score, n = 1, with_ties = FALSE) %>%
  arrange(desc(EA_score), desc(CAL1_score), domain, .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
    ca_ea_pair_score = EA_score * CAL1_score,
    pair_domain = factor(
      domain,
      levels = c("GA", "HU", "FI", "AH", "AG")
    )
  ) %>%
  filter(ID != 218 & ID != 73)

```


## make a little hist of this variable:
```{r}
ggplot(ca_hz, aes(x = ca_ea_pair_score, fill = pair_domain)) +
  geom_bar() +
  scale_fill_brewer(palette = "Paired", direction = -1, type = "qual") +
  labs(
    x = "Subsistence dependence × collective action score",
    y = "Number of societies",
    fill = "Dominant subsistence domain",
    title = "Spread of Collective Action within Dominant Subsistence Types"
  ) +
  theme_minimal()
```

## dealing with societies that have more than one dominant subsistence domain:
```{r}
ca_hz %>%
  filter(pair_coverage_count > 1) %>%
  select(ID)
```

SCCS = 100, 109

```{r}
ca_hz %>%
  filter(pair_coverage_count > 1) %>%
  select(
    ID,
    ca_ea_pair_score,
    pair_domain,
    matches("CAL1-R"),
    matches("^EA-")
  )
```


```{r}
ggplot(
  ca_hz %>%
    filter(pair_domain == "AG"),
  aes(
    x = ca_ea_pair_score,
    y = allhazards_H9a_sev_exposure_30
  )
) +
  geom_smooth(method = "loess",se = TRUE) +
  geom_jitter(
    aes(color = pair_domain),
    width = 0.1,
    height = 0,
    alpha = 0.7
  ) +
  labs(
    x = "Subsistence dependence × collective action score",
    y = "Food-destroying hazard severity weighted by frequency",
    color = "Dominant subsistence domain",
    title = "Hazard Severity Exposure by Subsistence Dependence–Collective Action Score"
  ) +
  theme_minimal()
```
## analyzing CA pair score with hazard vars
```{r}
data_clean <- ca_hz %>% filter(!is.na(ca_ea_pair_score), !is.na(allhazards_sev_exposure_30))  %>%
  mutate(
    AG_ind = ifelse(pair_domain == "AG", 1, 0),
    CA_ind = ifelse(!is.na(ca_ea_pair_score) & ca_ea_pair_score > 0, 1, 0)
  )

ca_lm <- lm(allhazards_sev_exposure_30 ~ ca_ea_pair_score*AG_ind, data = data_clean)

ca_ind_lm <- lm(allhazards_sev_exposure_30 ~ CA_ind*AG_ind, data = data_clean)

summary(ca_lm)
summary(ca_ind_lm)
```
The association between collective-action/subsistence pair score and hazard severity exposure differed by whether agriculture was the dominant subsistence domain. Among non-agriculturally dominant societies, higher pair scores were associated with lower hazard severity exposure, whereas among agricultural societies, higher pair scores were associated with higher hazard severity exposure. The interaction was statistically significant, suggesting that agricultural subsistence moderates this relationship.

```{r}
library(emmeans)

emtrends(
  ca_lm,
  ~ AG_ind,
  var = "ca_ea_pair_score"
)
```

Simple slopes showed that the association between collective-action/subsistence pair score and hazard severity exposure was positive among agricultural societies, b = 0.46, 95% CI [0.06, 0.87], but negative and not statistically significant among non-agricultural societies, b = -0.50, 95% CI [-1.20, 0.20].



Distribution of EA*CA score across agriculturally dominant societies
```{r}
d_ag_dist <- ca_hz %>%
  filter(pair_domain == "AG")

ggplot(d_ag_dist, aes(x = ca_ea_pair_score)) +
  geom_histogram(binwidth = 1, color = "white") +
  labs(
    x = "CA × EA pair score",
    y = "Count",
    title = "Distribution of CA–EA Pair Scores for Agriculture Societies"
  ) +
  theme_minimal()

```

